---
title: "PML_Project"
author: "Archer Lebron"
date: "September 24, 2017"
output: html_document
---
#Predicting barbell lift's exercise performance - How well are people doing it?
In this Coursera's Practical Machine Learning course project we will analyze exercise data (barbell lifts) of 6 participants to predict how well they did it.  In order to do this, we will analyze different variables, and predict the 'classe' variable that stands for the performance results of doing the exercise.  For this project Cross Validation, and classification prediction statistical method will be employed on a training set to predict 20 observations from a testing data set.

#Data Processing
##Getting the data
Load all the packages to be used in the analysis
```{r}
library(parallel); library(doParallel); library(caret); library(lattice)
library(ggplot2); library(rattle); library(randomForest); library(e1071)
```

Import the training and test data sets for PML course project
```{r}
pml_training <- read.csv("pml-training.csv", na.string=c("",".","NA"))
pml_testing <- read.csv("pml-testing.csv", na.string=c("", ".","NA"))
```


##Data cleaning
We'll delete columns that contain missing values in the training and test data sets
``` {r }
training <- pml_training[,colSums(is.na(pml_training)) == 0]
testing <- pml_testing[,colSums(is.na(pml_testing)) == 0]
```

Remove first seven predictors because they don't have an impact over the outcome `classe`.
```{r}
trainSubset <- training[,-c(1:7)]
testSubset <- testing[,-c(1:7)]
```


##Data splitting
Splitting the training data set into a new training data set (`newtrain`, 70%) to model fits for prediction on a new testing data set (`newtest`, 30%).
``` {r}
set.seed(8401)
inTrain <- createDataPartition(y=trainSubset$classe, p=0.70, list=FALSE)
newtrain <- trainSubset[inTrain,]
newtest <- trainSubset[-inTrain,]
```

#Fitting Models
For this project we will be applyin classification trees and random forests analysis to predict the outcome (`classe`).

##Random Forests
Create a k=10 k-fold cross validation for our predicting models. 
```{r}
#Configure parallel processing performance improvement
cluster <- makeCluster(detectCores() - 1)
registerDoParallel(cluster)
fitControl <- trainControl(method = "cv", number =10, allowParallel = TRUE)
```

Fit a random forest model using the train() function, with the `newtrain` data set and the `trainControl()` object we created for performance improvement.
``` {r}
rf_fit <- train(classe ~ ., method="rf", data=newtrain, trControl=fitControl)
stopCluster(cluster)
registerDoSEQ()
# Provide details for the random forest model
rf_fit
```

Predict outcome (**classe**) variable using the newtest data set with the Random Forest model fit.
``` {r}
rf_predict <- predict(rf_fit, newdata = newtest)
# Provide overall statistics of accuracy on the random forest predictions
rf_conf <- confusionMatrix(rf_predict, newtest$classe)
rf_conf
```

Looking at the Overall Statistics from the Random Forest model, we can see that the Accuracy is `r (rf_accuracy <- rf_conf$overall[1])`.

##Classification Trees
Using the same k-fold cross validation for the Random Forests fit, fit a Classification Trees model.
``` {r}
ct_fit <- train(classe ~., data=newtrain, method="rpart", trControl=fitControl)
ct_fit
```

Display Classification Trees visual graph using the fancyRpartPlot function
``` {r}
fancyRpartPlot(ct_fit$finalModel)
```

Predict outcome (**classe**) variable using the newtest data set with the Classification Trees model fit.
```{r}
ct_predict <- predict(ct_fit, newdata=newtest)
# Provide overall statistics of accuracy on the classification trees predictions
ct_conf <- confusionMatrix(ct_predict, newtest$classe)
ct_conf
```

Looking at the Overall Statistics from the Classification Trees model, we can see that the Accuracy is `r (ct_accuracy <- ct_conf$overall[1])`.

From this analysis we can state that the classification trees model does not preidct the outcome `classe` with high accuracy.

#Making prediction on the Testing data set
Now that we know that the Random Forests model provides a higher accuracy, we will apply this model to predict the oucome `classe` on the original testing data set.
``` {r}
predict_test <- predict(rf_fit, newdata = testing)
predict_test
```